\Large{Adversarial Variational Optimization of Non-Differentiable Simulators}

### 2017-06-26 tex/paper.tex
save

In many fields of science such as particle physics, climatology or population
genetics, computer simulators are used to describe complex processes that tie
parameters of an underlying theory to high dimensional observations. In most
cases, these implicit generative models~\citep{2016arXiv161003483M} are
specified as procedural implementations of forward stochastic processes that
generate data. Because it is usually computationally intractable, most simulators
do not provide a way to directly evaluate the likelihood function for a
given observation, thereby making inference difficult.
In addition, most scientific simulators are written using opaque low-level programming
languages, hence raising raising the bar for likelihood-free inference
algorithms relying e.g. on the simulator being a function with computable derivatives~\citep{2016arXiv160507826G}
or being a controllable probabilistic program~\citep{2016arXiv161009900L}.

@@

\subsection{Toy problem}

@@

As reviewed in \cite{2016arXiv161003483M}, likelihood-free inference is
intimately tied to a class of algorithms that can be framed as density
estimation-by-comparison. In most cases, these algorithms are formulated as an
iterative two-step process where the model distribution is first compared to the
true data distribution and then updated to make it more comparable to the
latter. Closest to our work are inference
algorithms~\citep{Neal:2007zz,gutmann2012noise,goodfellow2014generative,cranmer2015approximating,2016arXiv161110242D}
that rely on a classifier between generated and true data to estimate the
density ratio between the two distributions and make parameter updates.
\glnote{ABC + Classifier ABC.}
\glnote{Explain how this work is different.}

\glnote{Compare with gradient-free optimization.}

\glnote{Wood -> assume that the simulator is probabilistic program.}
\glnote{Storkey -> assume access to derivatives.}

>>>>

In many fields of science such as particle physics, climatology or population
genetics, computer simulators are used to describe complex processes that tie
parameters of an underlying theory to high dimensional observations. In most
cases, these implicit generative models~\citep{2016arXiv161003483M} are
specified as procedural implementations of forward stochastic processes that
generate data. Because it is usually computationally intractable, most simulators
do not provide a way to directly evaluate the likelihood function for a
given observation, thereby making inference difficult.
In addition, most scientific simulators are written using opaque low-level programming
languages, hence raising the bar for likelihood-free inference
algorithms relying e.g. on the simulator being a function with computable derivatives~\citep{2016arXiv160507826G}
or being a controllable probabilistic program~\citep{2016arXiv161009900L}.

@@

\subsection{Gaussian mixture model}

\subsection{Plinko}

@@

As reviewed in \cite{2016arXiv161003483M}, likelihood-free inference is
intimately tied to a class of algorithms that can be framed as density
estimation-by-comparison. In most cases, these inference algorithms are
formulated as an iterative two-step process where the model distribution is
first compared to the true data distribution and then updated to make it more
comparable to the latter. Closest to our work are procedures that rely on a
classifier to estimate the discrepancy between the true and the generated data
distributions. For example, \citep{gutmann2012noise} uses non linear logistic
regression for fitting unnormalized differentiable statistical models, while
\citep{goodfellow2014generative} exploits an adversarial neural network for
learning a differentiable implicit generative model. In the likelihood-free
setup, \citep{cranmer2015approximating,2016arXiv161110242D} estimate likelihood
ratios through supervised classification, which can in turn be used for
parameter inference in combination with a gradient-free optimization algorithm.
Similarly, \citep{gutmann2017likelihood} makes use of classification accuracy as
a summary statistics for approximate Bayesian computation.
\glnote{Explain what's new about the proposed work.}
